---
title: "p8106_mid_project"
author: "wq2151"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# read data
```{r}
# read data
original_data = read_csv("./data/2018_Financial_Data.csv") %>% 
  janitor::clean_names() %>% 
  rename(price2019 = "x2019_price_var_percent", 
         company = "x1")
```

# EDA and viz, get a rough idea about the data
```{r}
original_data %>% 
  mutate(class = as.factor(class)) %>% 
  group_by(class) %>% 
  summarise(number = n()) %>% 
  ggplot(aes(x = class, y = number)) + 
  geom_col(aes(fill = class)) +
  geom_text(aes(label = number), vjust = -0.1) + 
  labs(title = "Distribution of Stock Class", 
       x = "0 for BUY, 1 for SELL", y = "Number")

original_data %>% 
  group_by(sector) %>% 
  summarise(number = n()) %>% 
  ggplot(aes(x = reorder(sector, -number), y = number)) + 
  geom_col(aes(fill = sector)) +
  geom_text(aes(label = number), vjust = -0.1) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8), 
        legend.position = "none") + 
  labs(title = "Distribution of Company Sectors", 
       x = "Sector", y = "Number")
```

# check na values
```{r}
table(is.na(original_data))

# first we count the nan value in each row (for each company)
# 225 - 4 = 221 financial indicators
stock_na_count = original_data %>% 
  mutate(nan_row = rowSums(is.na(.)), 
         nan_row_ratio = nan_row/221)

stock_na_count %>% 
  ggplot(aes(x = reorder(company, -nan_row_ratio), y = nan_row_ratio)) +
  geom_col() + 
  theme(axis.text.x = element_blank()) + 
  labs(title = "Number of NA across rows", x = "Company", y = "NA ratio")
```

From the plot, we can see that there are lot of NA in each rows (company). Also, from the dataset we can see that some values are 0 where it couldn't be zero in the normal sense; for example, the R&D expense of GE is 0 (which is not correct). Therefore we assume that NA are the same as 0 in this dataset, produced by accounting/financial report errors. 

```{r}
# fill na with 0
stock = original_data %>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0))

# no na now
table(is.na(stock))
```

count zero values across rows
```{r}

```

